Generating Images From Captions\\ With Attention

### 2015-10-07 iclr-paper.tex
merged jimmys and elmans changes

\documentclass{article} % For LaTeX2e
\usepackage{iclr-style/iclr2016_conference,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{subfig}

\newcommand{\comm}[1]{}
\newcommand{\given}{\,|\,}
\newcommand{\expectation}{\mathbb{E}}
\newcommand{\kldiv}{\mathrm{D}_{\rm KL}}
\newcommand{\klBars}{\,\|\,}
\newcommand{\sigmoid}{\boldsymbol{\sigma}}
\newcommand{\hlang}{h^{lang}}
\newcommand{\hlangall}{\boldsymbol{h^{lang}}}
\newcommand{\hdec}{h^{dec}}
\newcommand{\henc}{h^{enc}}
\newcommand{\readop}{\mathit{read}}
\newcommand{\writeop}{\mathit{write}}
\newcommand{\encoder}{\mathit{LSTM}^{enc}}
\newcommand{\decoder}{\mathit{LSTM}^{dec}}
\newcommand{\canv}{c}
\newcommand{\lat}{z}
\newcommand{\Lat}{Z}
\newcommand{\numSamples}{L}
\newcommand{\sampleIdx}{l}
\newcommand{\LatSample}{\tilde{Z}}
\newcommand{\icaption}{{\bf{y}}}
\newcommand{\oimage}{{\bf{x}}}
\newcommand{\post}{Q}
\newcommand{\prior}{P}
\newcommand{\loss}{\mathcal{L}}
\newcommand{\lloss}{\mathcal{L}^{z}}
\newcommand{\rloss}{\mathcal{L}^{x}}

@@

There are two directions in learning a generative model of image and text.  One approach is to learn a text generative model conditioned on the images. Significant amount of recent works have been focusing on generating captions from images\citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc. The models take an image descriptor and generate unstructured texts through a recurrent decoder. By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description. 
Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation. Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image. Although, the interesting high dimensional natural images lay on a small manifold that is difficult to capture, the additional text description cues of a target image may simplify the learning problem by focusing on the conditional distribution.

@@

@@

We extended the architecture the DRAW network generative process to include additional input caption from the language model described in Sec. (\ref{sec:lang}). Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$. The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:

@@

where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $algin$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
\begin{align}
s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.
\end{align}

@@

\section{Experiments}
\subsection{MNIST With Captions}
As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly. Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background. The caption indicated the way digits were placed in the image and had the following template: \textit{the digit (1) is (2) the (3) of the digit (4) \textless eos\textgreater}. (1) and (4) were the digit numbers, (2) was either on or at, (3) was one of the adjectives top, bottom, left or right. For example, the generated images corresponding to the caption \textit{the digit seven is at the bottom of the digit five \textless eos\textgreater} as well as other captions are shown on~\ref{fig:fig1}. While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.

\begin{figure}[t]
\label{fig1}
\captionsetup{labelformat=empty}
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-4.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-6.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-9.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-13.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-17.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-22.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-26.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-27.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill

\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-4.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-6.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-9.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-13.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-17.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-22.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-26.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-27.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill

\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-4.png}
  \caption{\mbox{\textit{eos 0 of}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-6.png}
  \caption{\mbox{\textit{eos digit 0}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-9.png}
  \caption{\mbox{\textit{eos digit 0}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-12.png}
  \caption{\mbox{\textit{eos 0 of}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-17.png}
  \caption{\mbox{\textit{1 digit the}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-22.png}
  \caption{\mbox{\textit{1 digit the}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-25.png}
  \caption{\mbox{\textit{1 digit on}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-28.png}
  \caption{\mbox{\textit{digit the eos}}}
\endminipage\hfill

\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-4.png}
  \caption{\textit{left 8 4}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-6.png}
  \caption{\textit{left 8 4}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-9.png}
  \caption{\mbox{\textit{left digit 4}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-14.png}
  \caption{\mbox{\textit{eos 8 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-18.png}
  \caption{\mbox{\textit{eos 8 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-22.png}
  \caption{\mbox{\textit{8 digit the}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-25.png}
  \caption{\mbox{\textit{8 digit digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-28.png}
  \caption{\mbox{\textit{digit digit the}}}
\endminipage\hfill

\captionsetup{labelformat=empty}
\caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
\end{figure}

@@

Microsoft COCO \citep{mscoco} is a largest dataset of images annotated with captions consisting of 83k images. The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging. Since some of the images have more than five captions attached to them, for consistency with related work on caption generation we disregard extra captions.

In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with other related generative models.

\subsubsection{Analysis of Generated Images}
The main goal of this work is to learn a model that can understand the semantic meaning expressed in description of the image, such as the properties of objects, the relationships between them, etc. and then use that knowledge to generate relevant images. To verify that, we generated a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changed in the generated samples.

First, we wanted to see whether the model understood one of the simplest properties of any object, the color. We generated images of school buses with four different colors: yellow, red, green and blue. Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow. As you can see below, the model managed to generate images with an object that looked like a bus with the correct color.

\vspace{-0.3cm}
\begin{figure}[h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.22\textwidth]{figures/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.22\textwidth]{figures/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.22\textwidth]{figures/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.22\textwidth]{figures/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
%\caption{whole image caption}
\label{fig:label}
\vspace{-0.2cm}
\end{figure}

Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples. It is a somewhat harder task for a model than changing the color, due to larger number of changes that have to be made in the generated samples. Nevertheless as shown below, changing the grass type from dry to green as well as adding clouds in the caption has resulted in the appropriate changes.

\vspace{-0.3cm}
\begin{figure}[h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A very large commercial plane flying in the \underline{blue} skies.
]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in the \underline{cloudy} skies.
]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-cloudy-skies-sharp.png}}\quad
%
%
\end{center}
%\caption{whole image caption}
\label{fig:label}
\vspace{-0.8cm}
\end{figure}

\begin{figure}[h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A heard of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A heard of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
%
\end{center}
%\caption{whole image caption}
\label{fig:label}
\vspace{-0.2cm}
\end{figure}

Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we have found that the model made appropriate changes as long as some similar patter was present in the training set. However, it wasn't always the case when changing an object itself in the description. In cases, when objects did't have a noticeable differences in their properties, such as shape or color, the changes in the generated samples weren't very clearly seen. This could be attributed to the limitation of the model to grasp the detailed understanding of each object.

\vspace{-0.3cm}
\begin{figure}[h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[The \underline{decadent chocolate dessert} is on the table.
]{\includegraphics[width=0.48\textwidth]{figures/the-decadent-chocolate-dessert-is-on-the-table-2-sharp.png}}\quad
%
\subfloat[The \underline{bowl of banas} is on the table.
]{\includegraphics[width=0.48\textwidth]{figures/a-bowl-of-bananas-is-on-the-table-2-sharp.png}}\quad
%
%
\end{center}
%\caption{whole image caption}
\label{fig:label}
\vspace{-0.8cm}
\end{figure}

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A vintage photo of a \underline{cat}.
]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-cat-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{dog}.
]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-dog-sharp.png}}\quad
%
%
\end{center}
%\caption{whole image caption}
\label{fig:labelr}
\vspace{-0.2cm}
\end{figure}

@@

To be added soon.

\subsubsection{Comparison With Other Models}
The quantitative evaluation of generative models has been a subject of ambiguity within a machine learning community. Compared to reporting classification accuracies, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model. To avoid ambiguity, we report results on two different metrics as well as do qualitative comparison of generative models.

As shown on Figure blah, we have sampled the images from the same caption in four of the best models.

First, to compare the the peformance of our model with other generative models trained with variational objective, we report the Recall$@$K, namely the mean number of sentences for which the correct image is ranked within the top-K retrieved results, based on the lower bound of likelihood $\loss$. If you want to get a really good results build a model like Ryan did. Do deal away with easy images we take ratio with ... with mean caption. To deal with large computational complexity of looping over each image we only consider closest neightbors. The precision Recall curve is reported there...

Second,

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{iclr-style/iclr2016_conference,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{subfig}
\usepackage{tabulary}
\usepackage{soul}
\usepackage{color}
\usepackage{xcolor}

\captionsetup{font=small}

\newcommand{\comm}[1]{}
\newcommand{\given}{\,|\,}
\newcommand{\expectation}{\mathbb{E}}
\newcommand{\kldiv}{\mathrm{D}_{\rm KL}}
\newcommand{\klBars}{\,\|\,}
\newcommand{\sigmoid}{\boldsymbol{\sigma}}
\newcommand{\hlang}{h^{lang}}
\newcommand{\hlangall}{\boldsymbol{h^{lang}}}
\newcommand{\hdec}{h^{dec}}
\newcommand{\henc}{h^{enc}}
\newcommand{\readop}{\mathit{read}}
\newcommand{\writeop}{\mathit{write}}
\newcommand{\encoder}{\mathit{LSTM}^{enc}}
\newcommand{\decoder}{\mathit{LSTM}^{dec}}
\newcommand{\canv}{c}
\newcommand{\lat}{z}
\newcommand{\Lat}{Z}
\newcommand{\numSamples}{L}
\newcommand{\sampleIdx}{l}
\newcommand{\LatSample}{\tilde{Z}}
\newcommand{\icaption}{{\bf{y}}}
\newcommand{\oimage}{{\bf{x}}}
\newcommand{\post}{Q}
\newcommand{\prior}{P}
\newcommand{\loss}{\mathcal{L}}
\newcommand{\lloss}{\mathcal{L}^{z}}
\newcommand{\rloss}{\mathcal{L}^{x}}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}

\definecolor{0}{RGB}{247,251,255}
\definecolor{1}{RGB}{222,235,247}
\definecolor{2}{RGB}{198,219,239}
\definecolor{3}{RGB}{158,202,225}
\definecolor{4}{RGB}{107,174,214}
\definecolor{5}{RGB}{66,146,198}

\newcommand{\hlczero}[1]{\hlc[0]{#1}}
\newcommand{\hlcone}[1]{\hlc[1]{#1}}
\newcommand{\hlctwo}[1]{\hlc[2]{#1}}
\newcommand{\hlcthree}[1]{\hlc[3]{#1}}
\newcommand{\hlcfour}[1]{\hlc[4]{#1}}
\newcommand{\hlcfive}[1]{\hlc[5]{#1}}

@@

There are two directions in learning a generative model of image and text.  One approach is to learn a text generative model conditioned on the images. Significant amount of recent works has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc. The models take an image descriptor and generate unstructured texts through a recurrent decoder. By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description. 
Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation. Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image. Although, the interesting high dimensional natural images lay on a small manifold that is difficult to capture, the additional text description cues of a target image may simplify the learning problem by focusing on the conditional distribution.

@@

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\includegraphics[width=0.99\textwidth]{figures/alignDraw-cropped.pdf}\quad
%
\end{center}
\caption{The image of the proposed model}
\label{fig:figmodel}
\vspace{-0.3cm}
\end{figure}

@@

We extended the architecture the DRAW network generative process to include additional input caption from the language model described in Sec. (\ref{sec:lang}). Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$. The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:

@@

where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
\begin{align}
s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.
\end{align}

@@

\section{Experiments}
% \subsection{MNIST With Captions}
% As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly. Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background. The caption indicated the way digits were placed in the image and had the following template: \textit{the digit (1) is (2) the (3) of the digit (4) \textless eos\textgreater}. (1) and (4) were the digit numbers, (2) was either on or at, (3) was one of the adjectives top, bottom, left or right. For example, the generated images corresponding to the caption \textit{the digit seven is at the bottom of the digit five \textless eos\textgreater} as well as other captions are shown on~\ref{fig:fig1}. While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.

% \begin{figure}[t]
% \label{fig1}
% \captionsetup{labelformat=empty}
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-4.png}
%   \caption{\textit{is digit 5}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-6.png}
%   \caption{\textit{is digit 5}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-9.png}
%   \caption{\textit{is digit 5}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-13.png}
%   \caption{\textit{is digit 5}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-17.png}
%   \caption{\mbox{\textit{eos 7 digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-22.png}
%   \caption{\mbox{\textit{eos 7 digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-26.png}
%   \caption{\mbox{\textit{eos 7 digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/5-7-27.png}
%   \caption{\mbox{\textit{eos 7 digit}}}
% \endminipage\hfill

% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-4.png}
%   \caption{\mbox{\textit{eos digit 2}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-6.png}
%   \caption{\mbox{\textit{eos digit 2}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-9.png}
%   \caption{\mbox{\textit{eos digit 2}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-13.png}
%   \caption{\mbox{\textit{eos digit 2}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-17.png}
%   \caption{\mbox{\textit{of top digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-22.png}
%   \caption{\mbox{\textit{of top digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-26.png}
%   \caption{\mbox{\textit{of top digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/9-2-27.png}
%   \caption{\mbox{\textit{of top digit}}}
% \endminipage\hfill

% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-4.png}
%   \caption{\mbox{\textit{eos 0 of}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-6.png}
%   \caption{\mbox{\textit{eos digit 0}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-9.png}
%   \caption{\mbox{\textit{eos digit 0}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-12.png}
%   \caption{\mbox{\textit{eos 0 of}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-17.png}
%   \caption{\mbox{\textit{1 digit the}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-22.png}
%   \caption{\mbox{\textit{1 digit the}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-25.png}
%   \caption{\mbox{\textit{1 digit on}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/1-0-28.png}
%   \caption{\mbox{\textit{digit the eos}}}
% \endminipage\hfill

% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-4.png}
%   \caption{\textit{left 8 4}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-6.png}
%   \caption{\textit{left 8 4}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-9.png}
%   \caption{\mbox{\textit{left digit 4}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-14.png}
%   \caption{\mbox{\textit{eos 8 digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-18.png}
%   \caption{\mbox{\textit{eos 8 digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-22.png}
%   \caption{\mbox{\textit{8 digit the}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-25.png}
%   \caption{\mbox{\textit{8 digit digit}}}
% \endminipage\hfill
% \minipage{0.1\textwidth}
%   \includegraphics[width=\linewidth]{figures/4-8-28.png}
%   \caption{\mbox{\textit{digit digit the}}}
% \endminipage\hfill

% \captionsetup{labelformat=empty}
% \caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
% \end{figure}

@@

Microsoft COCO \citep{mscoco} is the largest dataset of images annotated with captions consisting of roughly 83k images. The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging. Since some of the images have more than five captions attached to them, for consistency with related work on caption generation we disregard extra captions.

In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with the performances of other related generative models.

\subsubsection{Analysis of Generated Images}
The main goal of this work is to learn a model that can understand the semantic meaning expressed in the descriptions of the images, such as the properties of objects, the relationships between them, etc. and then use that knowledge to generate relevant images. To verify that, we wrote a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.

First, we wanted to see whether the model understood one of the most basic properties of any object, the color. As shown in Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue. Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow. Despite that, the model managed to generate images of an object that is reminiscent of the bus pained with the relevant color.

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the color while keeping the caption fixed.}
\label{fig:genimages1}
\vspace{-0.3cm}
\end{figure}

Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples. Changing the background in images is a somewhat harder task for a model that changing the color, due to the larger visual area in images that is taken by the background. Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy as well as changing the grass type from dry to green resulted in the appropriate changes. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixelwise distance are displayed on top.}
\label{fig:genimages2}
\vspace{-0.3cm}
\end{figure}

Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar patter was present in the training set. However, it wasn't always the case when changing an object itself in the description. As you can see in Figure~\ref{fig:genimages3}, when objects didn't have a clear fine grained differences, such as different shape or color, the relevant changes in the generated samples weren't very clearly seen. This highlighted the limitation of the model to grasp the detailed understanding of each object.

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
%
\subfloat[\underline{A bowl of banas} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{cat}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{dog}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the object while keeping the caption fixed.}
\label{fig:genimages3}
\vspace{-0.3cm}
\end{figure}

@@

Unfortunately, we found that there was no real connection between the patches drawn on canvas and most attended words at each timestep. During the generation, the model mostly focused on several words that were represented the idea in the caption. The most attended words allowed us to interpret whether the model would make relevant changes.

\subsubsection{Comparison With Other Models}
The quantitative evaluation of generative models has been a subject of ambiguity in a machine learning community. Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model. To get a better comparisions between performances of generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
